Exploring probability distributions for bivariate temporal granularities

### 2019-12-09 main.Rmd
data structure change

Effective exploration or good visualization require good data structures. Commonly, simple sequences of time value pairs $(<t_0, v_0> \dots, <t_n, v_n>)$ are the basis of analysis and visualization. It is recognized that the initial approaches of just considering time as an ordinal dimension in visualisation are inadequate to capture characteristics of time-dependent information. There is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.  Moreover, one can use calendars based on application domain that define contextual system of granularities. Data can then be consolidated for different levels of granularity enabling statistical summaries of values along granularities. Since, we are interested in detection of previously unknown periodic behavior of data, it makes sense to support the detection of patterns by obtaining statistical summaries across cyclic time granularities. Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy. Instead, the big problem can be broken down into smaller pieces by focusing on two cyclic granularities at a time.

A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble], where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly. Since all cyclic granularities can be expressed in terms of the bottom granularity, which is the index of the tsibble, we consider the data structure in (Figure \@ref(fig:data-structure) for exploration of temporal data of this kind. This is extending the columns of tsibble by including the cyclic granularities.Now suppose we want to explore the measurement variable $v$ across two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}. Also, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$. Data subsets like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ such sets. These sets can form the basis for exploration and analysis of the measured variable across these bivariate cyclic granularities.

>>>>

Effective exploration or good visualization require good data structures. Commonly, simple sequences of time value pairs $(<t_0, v_0> \dots, <t_n, v_n>)$ are the basis of analysis and visualization. It is recognized that there is a crucial influence of linear vs cyclic time characteristics on the expressiveness of visualization and analysis.  Moreover, one can use calendars based on application domain that define contextual system of granularities. Data can then be consolidated for different levels of granularity enabling statistical summaries of values along granularities. Since, we are interested in detection of unknown periodic behavior of data, it makes sense to support the detection of patterns by obtaining statistical summaries across cyclic time granularities. Any attempt to encode all or many cyclic granularities at the same time to develop insights on periodicity might fail or become clumsy. Instead, the big problem is broken down into smaller pieces by focusing on two cyclic granularities at a time.

A recent tidy data structure to support exploration and modeling of temporal data is tsibble [@wang2019tsibble], where data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly. Since all cyclic granularities can be expressed in terms of the index set, we consider the data structure in (Figure \@ref(fig:data-structure) for exploration of temporal data of this kind. This is extending the columns of tsibble by including the cyclic granularities.Now suppose we want to explore the measurement variable $v$ across two cyclic granularities $C_1$ and $C_2$, such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_m$\}. Also, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = A_i$ and $C_2(s) = B_j$. Data subsets like <$A_i$, $B_j$, $v(s)$> can be obtained for all $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$  which will lead to $nm$ such sets. These sets can form the basis for exploration and analysis of the measured variable across these bivariate cyclic granularities.
